---
title: "Group#2 Homework# 3 - Logistic Regression"
author: "Group 2"
date: "3/29/2018"
output: html_document
---
#Introduction
This assignment explores, analyze and model a dataset containing information about crime in various neighborhoods in a major city. Each record has a response variable indicating whether or not the crime rate is above the median crime rate (1) or not (0). The crime dataset contains 18 variables and 466 observations. 
#Objective and Requirements
The objective is to build a binary logistic regression model on the training dataset to predict whether the neighborhood will be at risk for high crime levels. Classifications and probabilities for the evaluation dataset using the binary logistic regression model. 
#Approach
The team met to discuss this assignment and an approach to plan, complete the assignment. Each of the 5 team members was assigned tasks. The following tasks were assigned:
*Data Exploration
*Data Preparation
*Build Models
*Select Models

Github was used to manage the project. Using Github helped with version control and ensured each team member had access to the latest version of the project documentation. 

Slack was used to by the team to communicate during the project and for quick access to code and documentation.
```{r Libraries, echo=FALSE, message=FALSE, warning=FALSE}
library(psych)
library(dplyr)
library(ggplot2)
library(DataExplorer)
library(PerformanceAnalytics)
library(corrplot)
```

#Dataset
For reproducibility of the results, the data was loaded to and accessed from a Github repository. The age variable was rounded to a whole number.
```{r Load data, echo=FALSE, message=FALSE, warning=FALSE}
crime_trainData <- read.csv("https://raw.githubusercontent.com/indianspice/DATA621/master/Hw3/crime-training-data_modified%20(1).csv", header = 3)

crime_trainData$age <- round(crime_trainData[,6], digits = 0)

```

#Data Exploration
The following variables comprise the data set. The response variable (Target) is the variable of interest. The response variable is binary (0, 1) and identifies whether the crime rate is above the median crime rate. The remaining 12 variables are predictors. All variables are numeric. 
```{r Data Set Metadata, echo=FALSE, message=FALSE, warning=FALSE}
Variable_names <- c("zn","indus", "chas", "nox", "rm", "age", "dis", "rad", "tax",
                     "ptratio", "lstat", "medv", "traget")

Definitions <- c("proportion of residential land zoned for large lots (over 25000 square feet)", "proportion of non-retail business acres per suburb", "a dummy var. for whether the suburb borders the Charles River (1) or not (0)", "nitrogen oxides concentration (parts per 10 million)", " average number of rooms per dwelling", "proportion of owner-occupied units built prior to 1940", "weighted mean of distances to five Boston employment centers"," index of accessibility to radial highways", "full-value property-tax rate per $10,000", " pupil-teacher ratio by town", " lower status of the population (percent)", "median value of owner-occupied homes in $1000s", "whether the crime rate is above the median crime rate (1) or not (0)")

Variable_type <-c("Predictor", "Predictor", "Predictor", "Predictor", "Predictor", "Predictor", "Predictor", "Predictor", "Predictor", "Predictor", "Predictor", "Predictor", "Response")

Data_type <- c("Count", "Count", "Count", "Count", "Count", "Count", "Count", "Count", "Count", "Count", "Count", "Count", "Count")

df_crime_md <- cbind.data.frame (Variable_names, Definitions, Variable_type, Data_type)

colnames(df_crime_md) <- c("Variable Name", "Definition", "Variable Type", "Data Type") 
knitr::kable(df_crime_md)
```

Descriptive statistics were calculated to examine the basic featurs of the data. 
```{r EDA, echo=FALSE, message=FALSE, warning=FALSE}
#Use Describe Package to calculate Descriptive Statistic
(df_crime_des <- describe(crime_trainData, na.rm=TRUE, interp=FALSE, skew=TRUE, ranges=TRUE, trim=.1, type=3, check=TRUE, fast=FALSE, quant=c(.1,.25,.75,.90), IQR=TRUE))

#Calculate mean missing values per variable
crime_trainData %>% summarize_all(funs(sum(is.na(.)) / length(.)))

```

#zn
```{r zn, echo=FALSE, message=FALSE, warning=FALSE}
#Histogram
(h1 <- ggplot(crime_trainData, aes(x = zn)) + geom_histogram(colour = "black", fill = "light blue", binwidth = 4))

#Boxplot
(bp1 <- ggplot(crime_trainData, aes(x= " ", y = (zn))) + 
       stat_boxplot(geom ='errorbar') +
  geom_boxplot(fill = "light green", outlier.colour = "red", outlier.shape = 1))

#Examine outliers
get_outliers <-  function(x, n = 10) {
  
  v <- abs(x-mean(x,na.rm=TRUE)) > 3*sd(x,na.rm=TRUE)
  
  # capture all observations falling into outlier definition sort descending
  obs <- sort(unique(x[v]), decreasing = T)
  
  # handle cases where the number of observations is less than 
  # the parameter n to return for the top and bottom n values
  if (length(obs) < 2*n) {n <- floor(length(obs)/2)}
  
  hi <- obs[1:n]
  
  low <- obs[length(obs):(length(obs)-n +1)]
  
  # remove dupilcate entries from the lower bound outliers
  low <- setdiff(low, hi)

  return (list(Obs=obs, Hi=hi, Low=low))
  
}       

(o1 <- get_outliers(crime_trainData$zn))

```

#indus
```{r indus, echo=FALSE, message=FALSE, warning=FALSE}
#Histogram
(h2 <- ggplot(crime_trainData, aes(x = indus)) + geom_histogram(colour = "black", fill = "light blue", binwidth = 4))

#Boxplot
(bp2 <- ggplot(crime_trainData, aes(x= " ", y = (indus))) + 
       stat_boxplot(geom ='errorbar') +
  geom_boxplot(fill = "light green", outlier.colour = "red", outlier.shape = 1))

(o2 <- get_outliers(crime_trainData$indus))
```

#nox
```{r nox, echo=FALSE, message=FALSE, warning=FALSE}
#Histogram
(h4 <- ggplot(crime_trainData, aes(x = nox)) + geom_histogram(colour = "black", fill = "light blue", binwidth = 4))

#Boxplot
(bp4 <- ggplot(crime_trainData, aes(x= " ", y = (nox))) + 
       stat_boxplot(geom ='errorbar') +
  geom_boxplot(fill = "light green", outlier.colour = "red", outlier.shape = 1))

(o4 <- get_outliers(crime_trainData$nox))
```
#rm
```{r rm, echo=FALSE, message=FALSE, warning=FALSE}
#Histogram
(h5 <- ggplot(crime_trainData, aes(x = rm)) + geom_histogram(colour = "black", fill = "light blue", binwidth = 4))

#Boxplot
(bp5 <- ggplot(crime_trainData, aes(x= " ", y = (rm))) + 
       stat_boxplot(geom ='errorbar') +
  geom_boxplot(fill = "light green", outlier.colour = "red", outlier.shape = 1))

(o5 <- get_outliers(crime_trainData$rm))
```
#age
```{r age, echo=FALSE, message=FALSE, warning=FALSE}
#Histogram
(h5 <- ggplot(crime_trainData, aes(x = age)) + geom_histogram(colour = "black", fill = "light blue", binwidth = 4))

#Boxplot
(bp5 <- ggplot(crime_trainData, aes(x= " ", y = (age))) + 
       stat_boxplot(geom ='errorbar') +
  geom_boxplot(fill = "light green", outlier.colour = "red", outlier.shape = 1))

(o5 <- get_outliers(crime_trainData$age))
```

#dis
```{r dis, echo=FASLE, message=FALSE, warning=FALSE}
#Histogram
(h6 <- ggplot(crime_trainData, aes(x = dis)) + geom_histogram(colour = "black", fill = "light blue", binwidth = 4))

#Boxplot
(bp6 <- ggplot(crime_trainData, aes(x= " ", y = (dis))) + 
       stat_boxplot(geom ='errorbar') +
  geom_boxplot(fill = "light green", outlier.colour = "red", outlier.shape = 1))

(o6 <- get_outliers(crime_trainData$dis))
```

#rad
```{r rad, echo=FALSE, message=FALSE, warning=FALSE}
#Histogram
(h7 <- ggplot(crime_trainData, aes(x = rad)) + geom_histogram(colour = "black", fill = "light blue", binwidth = 4))

#Boxplot
(bp7 <- ggplot(crime_trainData, aes(x= " ", y = (rad))) + 
       stat_boxplot(geom ='errorbar') +
  geom_boxplot(fill = "light green", outlier.colour = "red", outlier.shape = 1))

(o7 <- get_outliers(crime_trainData$rad))
```

#tax
```{r tax, echo=FALSE, message=FALSE, warning=FALSE}
#Histogram
(h8 <- ggplot(crime_trainData, aes(x = tax)) + geom_histogram(colour = "black", fill = "light blue", binwidth = 4))

#Boxplot
(bp8 <- ggplot(crime_trainData, aes(x= " ", y = (tax))) + 
       stat_boxplot(geom ='errorbar') +
  geom_boxplot(fill = "light green", outlier.colour = "red", outlier.shape = 1))

(o8 <- get_outliers(crime_trainData$tax))
```
#ptratio
```{r ptratio, echo=FALSE, message=FALSE, warning=FALSE}
#Histogram
(h9 <- ggplot(crime_trainData, aes(x = ptratio)) + geom_histogram(colour = "black", fill = "light blue", binwidth = 4))

#Boxplot
(bp9 <- ggplot(crime_trainData, aes(x= " ", y = (ptratio))) + 
       stat_boxplot(geom ='errorbar') +
  geom_boxplot(fill = "light green", outlier.colour = "red", outlier.shape = 1))

(o9 <- get_outliers(crime_trainData$ptratio))
```
#lstat
```{r lstat, echo=FALSE, message=FALSE, warning=FALSE}
#Histogram
(h10 <- ggplot(crime_trainData, aes(x = lstat)) + geom_histogram(colour = "black", fill = "light blue", binwidth = 4))

#Boxplot
(bp10 <- ggplot(crime_trainData, aes(x= " ", y = (lstat))) + 
       stat_boxplot(geom ='errorbar') +
  geom_boxplot(fill = "light green", outlier.colour = "red", outlier.shape = 1))

(o10 <- get_outliers(crime_trainData$lstat))
```
#medv
```{r medv, echo=FALSE, message=FALSE, warning=FALSE}
#Histogram
(h11 <- ggplot(crime_trainData, aes(x = medv)) + geom_histogram(colour = "black", fill = "light blue", binwidth = 4))

#Boxplot
(bp11 <- ggplot(crime_trainData, aes(x= " ", y = (medv))) + 
       stat_boxplot(geom ='errorbar') +
  geom_boxplot(fill = "light green", outlier.colour = "red", outlier.shape = 1))

(o11 <- get_outliers(crime_trainData$medv))
```

# Correlation between variables
```{r correlation, echo=FALSE, message=FALSE, warning=FALSE}
chart.Correlation(crime_trainData[1:4])
chart.Correlation(crime_trainData[5:8])
chart.Correlation(crime_trainData[9:13])

crimeCorr <- cor(crime_trainData)
corrplot(crimeCorr, method = "circle")


```

